---
title: "creating a data frame"
author: "alfonso reyes"
date: "August 21, 2016"
output: html_document
---

## Creating a data frame from vectors

```{r}
patientID <- c(1, 2, 3, 4)
age <- c(25, 34, 28, 52)
diabetes <- c("Type1", "Type2", "Type1", "Type1")
status <- c("Poor", "Improved", "Excellent", "Poor")
patientdata <- data.frame(patientID, age, diabetes, status)
patientdata
```

### Specifying elements in a dataframe
```{r}
patientdata[1:2]
```

or:

```{r}
patientdata[c("diabetes", "status")]
```
which is similar to Python DataFrame:
df[["diabetes", "status"]]


Also, we can list a specific column with:
```{r}
patientdata$age
```


Cross tabulate diabetes type by status:

```{r}
table(patientdata$diabetes, patientdata$status)
```

## attach, dettach, with

```{r}
attach(mtcars)
  summary(mpg)
  plot(mpg, disp)
  plot(mpg, wt)
detach(mtcars)
```

or:

```{r}
with(mtcars, {
  summary(mpg, disp, wt)
  plot(mpg, disp)
  plot(mpg, wt)
})
```

If you need to create objects that will exist outside of the with() construct, use the special assignment operator <<- instead of the standard one (<-).

```{r}
with(mtcars, {
   nokeepstats <- summary(mpg)
   keepstats <<- summary(mpg)
})
# nokeepstats will produce error
keepstats
```




